---
output:
  word_document: default
  html_document: default
  pdf_document: default
---
##### Welcome to the R code of my master thesis: WILDFIRE RISK ASSESSMENT IN LIGURIA, comparison ##### of two geospatial modelling methods.

#### Be careful when you run the chunks entirely because some functions take a lot (really a lot) 
#### of computing time. Especially don't run the entire chunks "GWR", "GWR_prediction", "GRF" and 
#### "GRF_prediction".

# AUTHOR: M. Jérémie Fragnière, SUPERVISOR: Dr. Marj Tonini

```{r setup, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

```{r libraries, message=FALSE}
##### loading of the librairies #####
library(terra)
library(foreign)
library(readr) 
library(ggplot2)  
library(SpatialML)
library(GWmodel)
library(sp)
library(corrplot)
library(pROC)
library(randomForest)
library(caret)
library(classInt)
library(tidyr)
(.packages())
```


```{r histograms_fire_regime}
##### Histograms for the Ligurian fire regime #####

# Load the Dataset from Moris et al., 2014
Lig<-read.csv2("IB_Liguria_1986_2022.csv")

# Keep only the lines with data (it loads 16'328 empty lines I don't know why)
Lig<-Lig[1:36,]

# Find the non-numeric values and convert them for number of fires
non_numeric_values1 <- Lig$N_fires[is.na(as.numeric(Lig$N_fires))]
print(non_numeric_values1)
Lig$N_fires <- gsub("[^0-9.-]", "", Lig$N_fires)
Lig$N_fires<-as.numeric(Lig$N_fires)

# Find the non-numeric values and convert them for total burned area 
non_numeric_values2 <- Lig$BA_tot[is.na(as.numeric(Lig$BA_tot))]
print(non_numeric_values2)
Lig$BA_tot <- gsub("[^0-9.-]", "", Lig$BA_tot)
Lig$BA_tot<-as.numeric(Lig$BA_tot)


# Barplot of number of fires 
ggplot(Lig[11:36,], aes(x = factor(Year), y = N_fires)) +
  geom_bar(stat = "identity", fill = "lightblue",width=0.5) +
  labs(x = NULL, y = "Number of wildfires", title = "Number of wildfires (period 1997-2022)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) 

# Barplot of total burned area 
ggplot(Lig[11:36,], aes(x = factor(Year), y = BA_tot)) +
  geom_bar(stat = "identity", fill = "lightblue",width=0.5) +
  labs(x = NULL, y = "Burned area [ha]", title = "Burned area (period 1997-2022)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) 
```


```{r dataset_GWR, message=FALSE, warning=FALSE}
##### Computing the dataset for the GWR #####

#### Import raster of the independent variables at 100 meter resolution ####
Slope<-rast("slope.tif")
DEM<-rast("dem.tif")
DEM<-crop(DEM, Slope)
DEM<-project(DEM, Slope)
Northness<-rast("northness.tif")
Eastness<-rast("eastness.tif")
DistTo_Urban<-rast("dist_urban.tif")
DistTo_Roads<-rast("dist_roads.tif")
DistTo_Pathways<-rast("dist_pathways.tif")
Temperature<-rast("maenTemp_resampled.tif")
Precipitations<-rast("meanPrec_resampled.tif")
Vegetation<-as.factor(rast("VegAggr.tif"))

#### Plot the rasters ####
features <- c(DEM,Slope, Northness, Eastness, Vegetation, DistTo_Urban, DistTo_Roads, DistTo_Pathways, Temperature, Precipitations)
names(features)<-c("DEM","Slope", "Northness", "Eastness", "Vegetation", "DistTo_Urban", "DistTO_Roads", "DistTo_Pathways", "Temperature", "Precipitations")
features <- terra::mask(features, Slope)

#### Importation of the dependent variable of wildfire occurence #### 

# This dataset contains the presences (1) and absence (0) wildfires observations for a total of 10,000 points 
# Presence have been extracted randomly for burned areas (Polygon shapefile)
# Absence have been generated randomly for a same number of points 
WF_pa <- read.dbf("PresAbs_final_10M.dbf")

#### extraction of the values of the predictors at each location in the WF dataset. ####

# Shuffle the rows
WF_sh<-WF_pa [sample(nrow(WF_pa), nrow(WF_pa)), ]
summary(WF_sh)
class(WF_sh)

# Convert to vector
WF_vect=vect(WF_sh,geom=c("POINT_X","POINT_Y"),crs=crs(features))

# Extract values from features and add XY coordinates
WF_inputXY <-terra::extract(features, WF_vect, method="simple",  xy=TRUE)

# Add wildfire presence or absence variable 
WF_inputXY$WF <- WF_vect$grid_code

# Remove extra column of ID
WF_inputXY<-na.omit(WF_inputXY)
WF <- WF_inputXY[,-c(1)]

# Remove the two observations having "Lakes" for vegetation 
WF<-WF[WF$Vegetation != 10, ]

#### Encode vegetation ####

# dummify and encode the vegetation data
dmy <- dummyVars(" ~ .", data = WF)
WF <- data.frame(predict(dmy, newdata = WF))

# Reorganize dataframe
WF <- WF[,c(19,20,1:4,14:18,6:13,21)]
WF_GWR<-WF[,-c(10,12)]

# transform the dataframe into spatialpointdataframe for GWR
coordinates(WF_GWR)<-c("x","y")
```

```{r GWR}
##### Geographically weighted regression #####

# Calculate distance matrix
DM<-gw.dist(dp.locat=coordinates(WF_GWR))

# Find optimal bandwidth 
bw<- bw.gwr(WF~DEM+Slope+Northness+Eastness+DistTo_Urban+DistTO_Roads+DistTo_Pathways +Precipitations+Vegetation.21+Vegetation.22+Vegetation.23+Vegetation.32+Vegetation.33+Vegetation.34+Vegetation.37, data = WF_GWR, dMat=DM, approach = "AICc", kernel = "gaussian", adaptive = TRUE)
## Adaptive bandwidth (number of nearest neighbours): 48 AICc value: 7183.805       

#### Model selection procedure ####

# Definition of the dependent variable
DeVar <- "WF"

# Definition of the independent variables
InDeVars <- c("DEM","Slope","Northness","Eastness","DistTo_Urban","DistTO_Roads","DistTo_Pathways","Precipitations","Vegetation.21","Vegetation.22","Vegetation.23","Vegetation.32","Vegetation.33","Vegetation.34","Vegetation.37")

# Iterative computation of all the models
model.sel<-gwr.model.selection(DeVar, InDeVars, data=WF_GWR, bw=48, kernel="gaussian", approach="AIC", adaptive=TRUE,dMat=DM)

# Sorting and plotting of the models
sorted.models<-model.sort.gwr(model.sel,numVars=length(InDeVars),ruler.vector=model.sel[[2]][,2])
model.list<-sorted.models[[1]]
plot(sorted.models[[2]][,2],col="black",pch=20,lty=5,main="GWR model selection procedure",ylab="AICc",xlab="Model number",type="b")

# Compute gegraphically weighted regression 
gwr<-gwr.basic(WF~DEM+Slope+Northness+Eastness+DistTo_Urban+DistTO_Roads+DistTo_Pathways+Precipitations+Vegetation.22+Vegetation.23+Vegetation.32+Vegetation.33+Vegetation.34+Vegetation.37,data=WF_GWR,bw=48,dMat = DM, kernel="gaussian",adaptive=TRUE)

# Save the model 
saveRDS(gwr, "gwr.rds")

# Load the model 
gwr<-readRDS("gwr.rds")

#### Plotting results ####

# Extracting GWR outputs as dataframe 
gwrdf <- as.data.frame(gwr$SDF)

#### Local R2 ####

# plot
ggplot() +
  geom_point(data=gwrdf,aes(x=x,y=y.1,colour=Local_R2),size=1)+
  scale_color_gradientn(colors = c("white", "yellow","red", "blue"))+
  labs(fill = "R2", x = "X Coordinate", y = "Y Coordinate")+
  ggtitle("Local R2")

#### Local T values #### 

# Shrublands 
ggplot() +
  geom_point(data=gwrdf,aes(x=x,y=y.1,colour=Northness_TV),size=1)+
  scale_color_gradientn(colors = c("white", "yellow","red", "blue"))+
  labs(fill = "R2", x = "X Coordinate", y = "Y Coordinate")+
  ggtitle("Local T values of shrublands vegetation")

# Agro-forestry area
ggplot() +
  geom_point(data=gwrdf,aes(x=x,y=y.1,colour=Vegetation.22_TV),size=1)+
  scale_color_gradientn(colors = c("white", "yellow","red", "blue"))+
  labs(fill = "R2", x = "X Coordinate", y = "Y Coordinate")+
  ggtitle("Local T values of agro-forestry vegetation")

# Slope
ggplot() +
  geom_point(data=gwrdf,aes(x=x,y=y.1,colour=Slope_TV),size=1)+
  scale_color_gradientn(colors = c("white", "yellow","red", "blue"))+
  labs(fill = "R2", x = "X Coordinate", y = "Y Coordinate")+
  ggtitle("Local T values of Slope")
```

```{r GWR_prediction}
##### Geographically weighted regression as spatial predictor #####
#### WARNING: Be sure to run the chunk "GRF_prediction" below first to have the required objects. ####
#### The sets used for GRF where encoded directly to be sure to use the same separated data. ####

# Encode the training set
dmy_train <- dummyVars(" ~ .", data = WF_RF_train)
WF_GWR_train <- data.frame(predict(dmy_train, newdata = WF_RF_train))

# Encode the testing set
dmy_test <- dummyVars(" ~ .", data = WF_RF_test)
WF_GWR_test <- data.frame(predict(dmy_test, newdata = WF_RF_test))

# Reorganize the dataframe
WF_GWR_train <- WF_GWR_train[,c(1:6,16:19,9:15,20)]
WF_GWR_test <- WF_GWR_test[,c(1:6,16:19,9:15,20)]

# Rename the encoded columns 
colnames(WF_GWR_train)[11]<-"Vegetation.21"
colnames(WF_GWR_train)[12]<-"Vegetation.22"
colnames(WF_GWR_train)[13]<-"Vegetation.23"
colnames(WF_GWR_train)[14]<-"Vegetation.32"
colnames(WF_GWR_train)[15]<-"Vegetation.33"
colnames(WF_GWR_train)[16]<-"Vegetation.34"
colnames(WF_GWR_train)[17]<-"Vegetation.37"

colnames(WF_GWR_test)[11]<-"Vegetation.21"
colnames(WF_GWR_test)[12]<-"Vegetation.22"
colnames(WF_GWR_test)[13]<-"Vegetation.23"
colnames(WF_GWR_test)[14]<-"Vegetation.32"
colnames(WF_GWR_test)[15]<-"Vegetation.33"
colnames(WF_GWR_test)[16]<-"Vegetation.34"
colnames(WF_GWR_test)[17]<-"Vegetation.37"

# Change the dataframes into SpatialPointsDataFrame 
coordinates(WF_GWR_train)<-c("x","y")
coordinates(WF_GWR_test)<-c("x","y")

# Calculate the distance matrix
DM_train<-gw.dist(dp.locat=coordinates(WF_GWR_train))
DM_test<-gw.dist(dp.locat=coordinates(WF_GWR_train), rp.locat=coordinates(WF_GWR_test))

# Find the optimal bandwidth
bw_p<- bw.gwr(WF~DEM+Slope+Northness+Eastness+DistTo_Urban+DistTO_Roads+DistTo_Pathways +Precipitations+Vegetation.21+Vegetation.22+Vegetation.23+Vegetation.32+Vegetation.33+Vegetation.34+Vegetation.37, data = WF_GWR_train, dMat=DM_train, approach = "AICc", kernel = "gaussian", adaptive = TRUE)
## Adaptive bandwidth (number of nearest neighbours): 59 AICc value: 6087.883 

# Compute GWR as spatial predictor 
pred_gwr<- gwr.predict(WF~DEM+Slope+Northness+Eastness+DistTo_Urban+DistTO_Roads+DistTo_Pathways +Precipitations+Vegetation.21+Vegetation.22+Vegetation.23+Vegetation.32+Vegetation.33+Vegetation.34+Vegetation.37,data=WF_GWR_train,predictdata=WF_GWR_test,bw=59, kernel="gaussian",adaptive=TRUE,dMat1=DM_test,dMat2=DM_train)

saveRDS(pred_gwr,"pred_gwr.rds")

# Calculate RMSPE, MAPE, and z-scores 
RMSPE_gwr<-(mean((WF_GWR_test$WF - pred_gwr$SDF$prediction)^2))^0.5
MAPE_gwr <- mean(abs(WF_GWR_test$WF - pred_gwr$SDF$prediction))
zscore_gwr<-(WF_GWR_test$WF - pred_gwr$SDF$prediction)/(pred_gwr$SDF$prediction_var)^0.5
MeanZ_gwr<-mean(zscore_gwr)
SDZ_gwr<-(var(zscore_gwr))^0.5


# Plot the ROC curve to assess predictive performance 
roc_gwr<-roc(WF_GWR_test$WF,pred_gwr$SDF$prediction)
plot(1-roc_gwr$specificities,roc_gwr$sensitivities, type = 'l', col = 'blue',
 xlab = "False positive rate", ylab = "True positive rate",main="ROC curve for GWR predictions")

```

```{r dataset_GRF, message=FALSE, warning=FALSE}
##### Computing the dataset for the GRF #####
#### Import raster of the independent variables at 100 meter resolution ####
Slope<-rast("slope.tif")
DEM<-rast("dem.tif")
DEM<-crop(DEM, Slope)
DEM<-project(DEM, Slope)
Northness<-rast("northness.tif")
Eastness<-rast("eastness.tif")
DistTo_Urban<-rast("dist_urban.tif")
DistTo_Roads<-rast("dist_roads.tif")
DistTo_Pathways<-rast("dist_pathways.tif")
Temperature<-rast("maenTemp_resampled.tif")
Precipitations<-rast("meanPrec_resampled.tif")
Vegetation<-as.factor(rast("VegAggr.tif"))

#### Name the vegetation classes ####

# Extract current names
current_levels <- levels(Vegetation)[[1]]

# Create new classes names
new_class_names <- c(
 "10" = "Lakes",
 "11" = "Urban area",
 "21" = "Arable land, horticulture, permanent grassland",
 "22" = "Agro-forestry areas",
 "23" = "Nursery, ochards, olive grove, vineyards",
 "32" = "Chestnut and ilex groves, pine forests",
 "33" = "Beech and larch forests",
 "34" = "Mixed forest",
 "37" = "Shrublands")

# Ensure that new classes names match the existing levels
current_levels$VegAggr <- new_class_names[as.character(current_levels$ID)]
levels(Vegetation) <- current_levels
print(levels(Vegetation))

#### Plot the rasters ####
features <- c(DEM,Slope, Northness, Eastness, Vegetation, DistTo_Urban, DistTo_Roads, DistTo_Pathways, Temperature, Precipitations)
names(features)<-c("DEM","Slope", "Northness", "Eastness", "Vegetation", "DistTo_Urban", "DistTO_Roads", "DistTo_Pathways", "Temperature", "Precipitations")
features <- terra::mask(features, Slope)
plot(features)


#### Importation of the dependent variable of wildfire occurence #### 
# This dataset contains the presences (1) and absence (0) wildfires observations for a total of 10,000 points 
# Presence have been extracted randomly for burned areas (Polygon shapefile)
# Absence have been generated randomly for a same number of points 
WF_pa <- read.dbf("PresAbs_final_10M.dbf")

#### extraction of the values of the predictors at each location in the WF dataset. ####
# Shuffle the rows
WF_sh<-WF_pa [sample(nrow(WF_pa), nrow(WF_pa)), ]
summary(WF_sh)
class(WF_sh)
# Convert to vector (SpatVector:: point)
WF_vect=vect(WF_sh,geom=c("POINT_X","POINT_Y"),crs=crs(features))
# Extract values from features and add XY
WF_inputXY <-terra::extract(features, WF_vect, method="simple",  xy=TRUE)
# Add wildfire presence or absence variable 
WF_inputXY$WF <- WF_vect$grid_code
# Remove extra column of ID
WF_inputXY<-na.omit(WF_inputXY)
WF_RF <- WF_inputXY[,-c(1)]

# Reorganize dataframe
WF_RF <- WF_RF[,c(11,12,1:8,10,13)]
WF_RF <- WF_RF[WF_RF$Vegetation != "Lakes", ]

```


```{r GRF, message=FALSE, warning=FALSE}
##### Geographcially weighted random forest #####

# Set the x and y coordinates 
coords<-WF_RF[,1:2]

# Find optimal bandwidth
bwrf<-grf.bw(WF~DEM+Slope+Northness+Eastness+DistTo_Urban+DistTO_Roads+DistTo_Pathways
             +Precipitations+Vegetation,data=WF_RF, kernel="adaptive",coords=coords,bw.min=50,
             bw.max =65)
## Bandwidth: 63
## R2 of Local Model: 0.5978

# Compute geographically weighted random forest 
grf<-grf(WF~DEM+Slope+Northness+Eastness+DistTo_Urban+DistTO_Roads+DistTo_Pathways
        +Precipitations+Vegetation,dframe =WF_RF,kernel="adaptive",bw=63,
        coords=coords)

# Save the model 
saveRDS(grf, "grf.rds")

# Load the model 
grf<-readRDS("grf.rds")

#### Plotting results ####

#### Global variable importance ####

# Extract values 
var_imp<-data.frame(
  Variable = c("DEM","Slope","Northness","Eastness","DistTo_Urban","DistTO_Roads","DistTo_Pathways","Precipitations","Vegetation"),
  Importance = c(264.6651,255.8067,229.7007,187.3525,201.2740,167.6852,185.0279,360.7378,447.1773)
)

# Set color for the three most important features
var_imp$Color <- ifelse(var_imp$Importance >= sort(var_imp$Importance, decreasing = TRUE)[3], "orange", "skyblue")

# plot
ggplot(data = var_imp, aes(x = Variable, y = Importance, fill = Color)) +
  geom_bar(stat = "identity") +
  scale_fill_identity() +
  labs(title = "Variable Importance Plot", x = "Variable", y = "Importance") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#### Local R2 #### 

# Extract values
Rsq<-grf$LGofFit$LM_Rsq100

# Add coordinates 
Rsq_XY<-as.data.frame(cbind(Rsq,grf$Locations$x,grf$Locations$y))
colnames(Rsq_XY)[2]<- "X"
colnames(Rsq_XY)[3]<- "Y"

# Add GWR R2
Rsq_XY<-cbind(Rsq_XY,gwr$SDF$Local_R2)

# Reshape the data into long format 
Rsq_XY_long <- Rsq_XY %>%
  select(Rsq, Local_R2 = `gwr$SDF$Local_R2`) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value")

# Create boxplot for comparing local R2 distribution of two models 
ggplot(Rsq_XY_long, aes(x = Variable, y = Value, fill = Variable)) +
  geom_boxplot() +
  scale_fill_manual(
    values = c("skyblue", "orange"),         
    labels = c("GWR", "GRF")             
  ) +
  labs(title = "Distribution of Local R²",
       x = "Variable",
       y = "Value",
       fill = "Model") +                  
  theme_minimal()

# plot
ggplot () +
  geom_point(data = Rsq_XY, aes(x = X, y = Y, colour = Rsq), size = 1)+ 
  scale_color_gradientn(colors = c("white","yellow" ,"red", "blue"))+
  labs(title = "R2", x = "X Coordinate", y = "Y Coordinate")+
  ggtitle("Local R2")

#### Variables Local Importance ####

# Extract values 
grf_LI<-grf$Local.Variable.Importance

# Add coordinates 
grf_LI_XY<-cbind(grf_LI,grf$Locations$x,grf$Locations$y)
colnames(grf_LI_XY)[10]<- "X"
colnames(grf_LI_XY)[11]<- "Y"

# Define breaks with fisher method for feature importance plots
vege_vect <- as.vector(grf_LI_XY$Vegetation)
vege_brk<-(classIntervals(vege_vect, n=5, style = "fisher"))
vege_brkInt<-round(vege_brk$brks, digits=2)
print(vege_brkInt)

prec_vect <- as.vector(grf_LI_XY$Precipitations)
prec_brk<-(classIntervals(prec_vect, n=5, style = "fisher"))
prec_brkInt<-round(prec_brk$brks, digits=2)
print(prec_brkInt)

DEM_vect <- as.vector(grf_LI_XY$DEM)
DEM_brk<-(classIntervals(DEM_vect, n=5, style = "fisher"))
DEM_brkInt<-round(DEM_brk$brks, digits=2)
print(DEM_brkInt)

# Plot for Vegetation
ggplot() +
    geom_point(data = grf_LI_XY, aes(x = X, y = Y, colour = Vegetation), size = 1) +
    scale_color_gradientn(colors = c("white", "orange","red", "blue"),  
                        breaks = c(0.00, 0.33, 0.88, 1.76, 3.28, 7.28), 
                        labels=c(0.00, 0.33, 0.88, 1.76, 3.28, 7.28)) +
        labs( x = "X Coordinate", y = "Y Coordinate")+
    ggtitle("Local average importance of vegetation")

# Plot for Precipitations 
ggplot() +
    geom_point(data = grf_LI_XY, aes(x = X, y = Y, colour = Precipitations), size = 1) +
    scale_color_gradientn(colors = c("white", "orange","red", "blue"),
                         breaks = c(0.00, 0.60, 1.36, 2.30, 3.64, 7.21), 
                        labels=c(0.00, 0.60, 1.36, 2.30, 3.64, 7.21)) +
        labs( x = "X Coordinate", y = "Y Coordinate")+
    ggtitle("Local average importance of precipitations")

# Plot for Temperature 
ggplot() +
    geom_point(data = grf_LI_XY, aes(x = X, y = Y, colour = DEM), size = 1) +
    scale_color_gradientn(colors = c("white", "orange","red", "blue"),
                        breaks = c(0.00, 0.62, 1.45, 2.55, 3.88, 6.67), 
                        labels=c(0.00, 0.62, 1.45, 2.55, 3.88, 6.67)) +
        labs( x = "X Coordinate", y = "Y Coordinate")+
    ggtitle("Local average importance of DEM")

```

```{r local_RF package}
##### Local Random Forest ##### 

# Compute local model
rf<-randomForest(WF~DEM+Slope+Northness+Eastness+DistTo_Urban+DistTO_Roads+DistTo_Pathways
                 +Precipitations+Vegetation,data=WF_RF,importance=TRUE)

# Exctract variables importance to make sure that result is the same as GRF
rf_imp<-importance(rf)
## Result is the same as GRF

# Calculate R2
rf_R2 = 1 - sum((rf$y-rf$predicted)^2)/sum((rf$y-mean(rf$y))^2)
## 0.41

# Partial dependence plot for vegetation. 
partialPlot(rf, WF_RF, x.var = Precipitations,rug=TRUE, xlab= "Vegetation classes", main = "Importance of vegetation classes", ylab = "Partial dependence")

#### Predictions ####

#Train new model
rf_p<-rf<-randomForest(WF~DEM+Slope+Northness+Eastness+DistTo_Urban+DistTO_Roads+DistTo_Pathways
                 +Precipitations+Vegetation,data=WF_RF_train,importance=TRUE)

# Exctract variables importance to make sure that result is the same as GRF
rf_imp_p<-importance(rf_p)
## Result is the same as GRF

# Make prediction with new local model 
pred_rf<-predict(object = rf_p, newdata = WF_RF_test, type="response",predict.all=TRUE)

# Compute variance of the predictions 
pred_rf_var<-apply(pred_rf$individual,1,var)

# Calculate RMSPE, MAPE, and z-scores 
RMSPE_rf<-(mean((WF_RF_test$WF - pred_rf$aggregate)^2))^0.5
MAPE_rf <- mean(abs(WF_RF_test$WF - pred_rf$aggregate))
zscore_rf<-(WF_RF_test$WF - pred_rf$aggregate)/(pred_rf_var)^0.5
MeanZ_rf<-mean(zscore_rf)
SDZ_rf<-(var(zscore_rf))^0.5


# Plot the ROC curve to assess predictive performance 
roc_rf<-roc(WF_RF_test$WF,pred_rf$aggregate)
plot(1-roc_rf$specificities,roc_rf$sensitivities, type = 'l', col = 'blue',
 xlab = "False positive rate", ylab = "True positive rate",main="ROC curve for RF global model predictions")


```

```{r GRF_prediction}
##### Geographically weighted random forest as spatial predictor #####

# Split the input dataset into training (80%) and testing (20%)
n <- nrow (WF_RF)
n_train <- round(0.80 * n) 
train_indices <- sample(1:n, n_train)
WF_RF_train <- WF_RF[train_indices, ]  
WF_RF_test <- WF_RF[-train_indices, ]

# Set x and y coordinates 
coords_p<-WF_RF_train[,1:2]

# Find optimal bandwidth 
bwrf_p<-grf.bw(WF~DEM+Slope+Northness+Eastness+DistTo_Urban+DistTO_Roads+DistTo_Pathways
        +Precipitations+Vegetation,
        data=WF_RF_train, kernel="adaptive",coords=coords_p, bw.min=50, bw.max = 65)
## Best Bandwidth (Based on the Local Model): 59
## R2 of Local Model: 0.574181709775723

# Compute new model 
grf_p<-grf(WF~DEM+Slope+Northness+Eastness+DistTo_Urban+DistTO_Roads+DistTo_Pathways
        +Precipitations+Vegetation,dframe =WF_RF_train,kernel="adaptive",bw=59,
        coords=coords_p)

# Compute the predictions
pred_grf<-predict.grf(grf_p,WF_RF_test,x.var.name="x",y.var.name="y",predict.all=TRUE)

# Calculate RMSPE, MAPE, and z-scores 
RMSPE_grf<-(mean((WF_RF_test$WF - pred_grf)^2))^0.5
MAPE_grf <- mean(abs(WF_RF_test$WF - pred_grf))
zscore_grf<-(WF_RF_test$WF - pred_rf$aggregate)/(pred_rf_var)^0.5
MeanZ_grf<-mean(zscore_rf)
SDZ_grf<-(var(zscore_rf))^0.5


# Plot the ROC curve to assess predictive performance 
roc_grf<-roc(WF_RF_test$WF,pred_grf)
plot(1-roc_grf$specificities,roc_grf$sensitivities, type = 'l', col = 'blue',
 xlab = "False positive rate", ylab = "True positive rate",main="ROC curve for GRF predictions")




```

